### 2021-03-03 02_abstract/abstract_english.tex
Added English abstract and first part of acknowledgements

A precise quantification of climate change is crucial in order to assess
optimal mitigation and adaptation strategies. Valuable tools in this context
are \acp{ESM}, which are state-of-the-art climate models that allow numerical
simulations of the complex physical, biological and chemical processes of the
Earth system. Of particular relevance are projections of the future climate,
which enable an assessment of future climate change. Due to the chaotic nature
of the climate system, unknowns in future emission pathways and uncertainties
in the climate models, projections of the future climate are associated with
large uncertainties. In this thesis, we analyze future climate projections from
\acp{ESM} participating in the \ac{CMIP} and reduce corresponding uncertainties
with observations.

>>>>

A precise quantification of climate change is crucial in order to assess
optimal mitigation and adaptation strategies. Valuable tools in this context
are \acfpAbstract{ESM}, which are state-of-the-art climate models that allow
numerical simulations of the complex physical, biological and chemical
processes of the Earth system. Of particular relevance are projections of the
future climate, which enable an assessment of future climate change. Due to the
chaotic nature of the climate system, unknowns in future emission pathways and
uncertainties in the climate models, projections of the future climate are
associated with large uncertainties. In this thesis, we analyze future climate
projections from \acsp{ESM} participating in the \acfAbstract{CMIP} and reduce
corresponding uncertainties with observations.

In a first step, the climate sensitivity (\ie{} the temperature response of the
climate system to an external forcing) in the latest generation of \acsp{ESM}
from \acs{CMIP}6 is evaluated. For the \acfAbstract{ECS}, the equilibrium
temperature response that follows a doubling of the atmospheric
\acfAbstract{CO2} concentration, we find a \acfAbstract{MMM} and multi-model
range of $3.74 \unit{K}$ and $\rangeunit{1.8}{5.6}{K}$, respectively. These
values are higher than in any previous \acs{CMIP} ensemble before. Moreover, a
third of the analyzed \acs{CMIP}6 models exceeds the upper bound of the likely
\acs{ECS} range of $\rangeunit{1.5}{4.5}{K}$ assessed by the Fifth \acl{AR}
(\acs{AR}5) of the \acl{IPCC} from 2013. Similarly, the transient response of
the climate system to a doubling of \acs{CO2}, also known as \acfAbstract{TCR},
shows an inter-model range of $\rangeunit{1.3}{3.0}{K}$, whose upper bound is
again higher than the likely range assessed in \acs{AR}5 of
$\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased climate
sensitivity in many \acs{CMIP}6 models is a change in the microphysical
representation of mixed-phase clouds above the Southern Ocean, which reduces
the strong negative shortwave cloud phase change feedback in this region that
is present in climate models from previous \acs{CMIP} generations.

To reduce uncertainties in \acs{ECS} projected by the \acs{CMIP}6 models, we
systematically analyze eleven published emergent constraints on \acs{ECS} that
have mostly been derived from models participating in \acs{CMIP}5, the
predecessor generation of \acs{CMIP}6. Emergent constraints are potentially
promising approaches to reduce uncertainties in climate model projections by
combining observations and output from \acsp{ESM}. The focus of this study is
on testing if these emergent constraints hold for \acsp{ESM} participating
\acs{CMIP}6. Since none of the emergent constraints considered here have been
derived on the \acs{CMIP}6 ensemble, the \acs{CMIP}6 models can be used for
cross-checking of the emergent constraints on a new model ensemble. The
application of the emergent constraints to \acs{CMIP}6 data shows a decrease in
skill and statistical significance of the emergent relationship for nearly all
constraints, with this decrease being large in many cases. Consequently, the
size of the constrained \acs{ECS} ranges ($66 \unit{\%}$ confidence intervals)
widens by $51 \unit{\%}$ on average in \acs{CMIP}6 compared to \acs{CMIP}5.
This is likely related to the increased multi-model spread of \acs{ECS} in
\acs{CMIP}6, but may in some cases also be due to spurious statistical
relationships or a too small number of models in the ensemble that the emergent
constraint was originally derived from. The corresponding best estimates given
by the emergent constraints also increase from \acs{CMIP}5 to \acs{CMIP}6 by
$12 \unit{\%}$ on average. This can be at least partly explained by the
increased number of high-\acs{ECS} models in \acs{CMIP}6 without a
corresponding change in the constraint predictors, suggesting the emergence of
new feedback processes rather than changes in strength of those previously
dominant. Our results support previous studies concluding that emergent
constraints should be based on an independently verifiable physical mechanism
and that process-based emergent constraints on \acs{ECS} should rather be
thought of as constraints for the process or feedback they are actually
targeting.

To overcome these issues of single-process-oriented emergent constraints, we
introduce an alternative approach based on \acfAbstract{ML}. Since this new
technique relies on a large number of data points in order to train the
\acs{ML} algorithm, the scalar climate sensitivity expressed as \acs{ECS} or
\acs{TCR} is not an appropriate target variable. As an alternative, we study
projections of the future \acfAbstract{GPP}, which is the largest flux of the
terrestrial carbon uptake and slows down global warming by removing \acs{CO2}
from the atmosphere. In this analysis, we combine an existing emergent
constraint on \acs{CO2} fertilization with a \acs{ML} approach to constrain the
spatial variations of multi-model \acs{GPP} projections. In the first step of
our two-step approach, we use observed changes in the \acs{CO2} seasonal cycle
at \acl{KUM} to constrain the global mean \acs{GPP} at the end of the \nth{21}
century (\range{2091}{2100}) in \acl{RCP} 8.5 simulations with \acsp{ESM}
participating in \acs{CMIP}5 to $\pmrangeunit{171}{12}{\GtCyr{}}$, compared to
the unconstrained model range $\rangeunit{156}{247}{\GtCyr{}}$. In a second
step, we use a \acs{ML} model to constrain gridded future absolute \acs{GPP}
and gridded fractional \acs{GPP} change in two independent approaches. For
this, observational data is fed into the \acs{ML} algorithm that has been
trained on \acs{CMIP}5 data to learn relationships between present-day
physically-relevant diagnostics and the target variable. In a
leave-one-model-out \acl{CV} approach, the \acs{ML} model shows superior
performance to the \acs{CMIP}5 \acs{MMM}. Our approach predicts an increased
\acs{GPP} change in northern high latitudes compared to regions closer to the
equator.

### 2021-03-03 ch07_conclusion/conclusion.tex
Added English abstract and first part of acknowledgements

In a first analysis, we evaluate the climate sensitivity metrics \ac{ECS} and
\ac{TCR} for the latest generation of \acp{ESM} from \acs{CMIP}6. This work,
which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and already
published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
\ac{ECS}, we find a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$, which
is higher on the upper and lower end than any model range from previous
\ac{CMIP} generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6
\ac{MMM} of \ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in
\acs{CMIP}6 versus $3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the
upper bound of the \acs{CMIP}6 models range is considerably larger than in the
assessed range of $\rangeunit{1.5}{4.5}{K}$ given by the latest published
\ac{AR} of the \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper
bound of $4.5 \unit{K}$ is exceeded by about a third of the \acs{CMIP}6 models,
with many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the
model range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
\acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased
climate sensitivity in many \acs{CMIP}6 models is a change in the microphysical
representation of mixed-phase clouds above the Southern Ocean. For one thing,
this change leads to an improved shortwave \ac{CRE} in these models when
compared to observations. However, this change also substantially reduces the
strong negative shortwave cloud feedback over the Southern Ocean that is
present in previous \ac{CMIP} generations and that results from a cloud phase
change from ice clouds in the present-day to liquid clouds in the future. In
the affected \acs{CMIP}6 models, this cloud phase change is not possible
anymore since the models predominantly simulate liquid clouds over the
present-day Southern Ocean.

@@

To overcome these issues of single-process-oriented emergent constraints,
\cref{ch:06:paper_gpp} introduces an alternative approach based on \ac{ML}.
This work is already published in \textcite{Schlund2020}. Since the new
technique relies on a large number of data points in order to train the \ac{ML}
algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is
not an appropriate target variable. As an alternative, we study projections of
the future \ac{GPP}, which is the largest flux of the terrestrial carbon uptake
and slows down global warming by removing \ac{CO2} from the atmosphere. In the
first step of our new two-step approach, we constrain the global mean \ac{GPP}
at the end of the \nth{21} century in the \acs{CMIP}5 \acs{RCP}8.5 scenario to
$\pmrangeunit{171}{12}{\GtCyr{}}$ using a published emergent constraint by
\textcite{Wenzel2016}. This first step corrects the \acs{CMIP}5 models'
\ac{GPP} response to \ac{CO2}, which shows a range of
$\rangeunit{156}{248}{\GtCyr{}}$ in the raw multi-model ensemble. In the second
step, a \ac{ML}-based climate model weighting approach is used to further
constrain the gridded \ac{GPP} based on present-day predictors that are
relevant for the simulation of \ac{GPP} in the \acp{ESM}. Our \ac{ML} approach
is mathematically similar to the \ac{MDER} approach \autocite{Karpechko2013,
  Senftleben2020, Wenzel2016a}, but additionally considers multi-dimensional
(gridded) target variables and non-linear relationships between the predictors
and the target variable. We establish a relationship between process‐oriented
predictors and future projections of \ac{GPP} and then utilize this to project
today's observed conditions into the future. The prediction phase of our method
can be interpreted as an implicit performance weighting. However, due to the
complex structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible
to extract specific values for the individual weights. We consider two target
variables: the gridded monthly climatologies of absolute \ac{GPP}
(\range{2091}{2100}) and the gridded fractional \ac{GPP} change over the
\nth{21} century (2100 versus 2000). The latter quantity shows an increased
\acs{GPP} change in the high latitudes compared to regions closer to the
equator. The results of both approaches are consistent with each other and with
the global constraint of the first step. We validate our approach by comparing
it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
\ac{LASSO} model) in a leave‐one‐model‐out \ac{CV} setup. Compared to \ac{MMM}
(\ac{LASSO}), we find a reduction of the resulting mean \ac{RMSEP} of up to $48
\unit{\%}$ ($3 \unit{\%}$) when using our \ac{ML} approach. Moreover, the
evaluation of the global and local feature importance allows us to get further
insights into the \ac{ML} model. For the first target variable (absolute
\ac{GPP}), historical \ac{GPP} is by far the most important predictor, which
can be explained with a correction of the historical bias in \ac{GPP} by our
approach. For the second variable, \ac{T} and \ac{LAI} are the dominant
features. This study directly addresses \cref{enum:01:question_3}
(\enquote{\emph{\KeyScienceQuestionThree{}}}), which we can confidently answer
with \enquote{yes} based on the results we find.

>>>>

In a first analysis, we evaluate the climate sensitivity metrics \ac{ECS} and
\ac{TCR} for the latest generation of \acp{ESM} from \acs{CMIP}6. This work,
which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and already
published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
\ac{ECS}, we find a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$, which
is higher on the upper and lower end than any model range from previous
\ac{CMIP} generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6
\ac{MMM} of \ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in
\acs{CMIP}6 versus $3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the
upper bound of the \acs{CMIP}6 models range is considerably larger than in the
assessed range of $\rangeunit{1.5}{4.5}{K}$ given by the latest published
\ac{AR} of the \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper
bound of $4.5 \unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with
many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model
range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
\acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased
climate sensitivity in many \acs{CMIP}6 models is a change in the microphysical
representation of mixed-phase clouds above the Southern Ocean. For one thing,
this change leads to an improved shortwave \ac{CRE} in these models when
compared to observations. However, this change also substantially reduces the
strong negative shortwave cloud feedback over the Southern Ocean that is
present in previous \ac{CMIP} generations and that results from a cloud phase
change from ice clouds in the present-day to liquid clouds in the future. In
the affected \acs{CMIP}6 models, this cloud phase change is not possible
anymore since the models predominantly simulate liquid clouds over the
present-day Southern Ocean.

@@

To overcome these issues of single-process-oriented emergent constraints,
\cref{ch:06:paper_gpp} introduces an alternative approach based on \ac{ML}.
This work is already published in \textcite{Schlund2020}. Since the new
technique relies on a large number of data points in order to train the \ac{ML}
algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is
not an appropriate target variable. As an alternative, we study projections of
the future \ac{GPP}, which is the largest flux of the terrestrial carbon uptake
and slows down global warming by removing \ac{CO2} from the atmosphere. In the
first step of our new two-step approach, we constrain the global mean \ac{GPP}
at the end of the \nth{21} century in the \acs{CMIP}5 \acs{RCP}8.5 scenario to
$\pmrangeunit{171}{12}{\GtCyr{}}$ using a published emergent constraint by
\textcite{Wenzel2016}. This first step corrects the \acs{CMIP}5 models'
\ac{GPP} response to \ac{CO2}, which shows a range of
$\rangeunit{156}{247}{\GtCyr{}}$ in the raw multi-model ensemble. In the second
step, a \ac{ML}-based climate model weighting approach is used to further
constrain the gridded \ac{GPP} based on present-day predictors that are
relevant for the simulation of \ac{GPP} in the \acp{ESM}. Our \ac{ML} approach
is mathematically similar to the \ac{MDER} approach \autocite{Karpechko2013,
  Senftleben2020, Wenzel2016a}, but additionally considers multi-dimensional
(gridded) target variables and non-linear relationships between the predictors
and the target variable. We establish a relationship between process‐oriented
predictors and future projections of \ac{GPP} and then utilize this to project
today's observed conditions into the future. The prediction phase of our method
can be interpreted as an implicit performance weighting. However, due to the
complex structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible
to extract specific values for the individual weights. We consider two target
variables: the gridded monthly climatologies of absolute \ac{GPP}
(\range{2091}{2100}) and the gridded fractional \ac{GPP} change over the
\nth{21} century (2100 versus 2000). The latter quantity shows an increased
\acs{GPP} change in the high latitudes compared to regions closer to the
equator. The results of both approaches are consistent with each other and with
the global constraint of the first step. We validate our approach by comparing
it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
\ac{LASSO} model) in a leave‐one‐model‐out \ac{CV} setup. Compared to \ac{MMM}
(\ac{LASSO}), we find a reduction of the resulting mean \ac{RMSEP} of up to $48
\unit{\%}$ ($3 \unit{\%}$) when using our \ac{ML} approach. Moreover, the
evaluation of the global and local feature importance allows us to get further
insights into the \ac{ML} model. For the first target variable (absolute
\ac{GPP}), historical \ac{GPP} is by far the most important predictor, which
can be explained with a correction of the historical bias in \ac{GPP} by our
approach. For the second variable, \ac{T} and \ac{LAI} are the dominant
features. This study directly addresses \cref{enum:01:question_3}
(\enquote{\emph{\KeyScienceQuestionThree{}}}), which we can confidently answer
with \enquote{yes} based on the results we find.

### 2021-03-03 end_acknowledgments/acknowledgments.tex
Cleaned tex files and added text to chapter 4
Added English abstract and first part of acknowledgements

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                          Declaration of Authorship                          %
%=============================================================================%

\chapter{Declaration of Authorship}

I assure that this thesis is a result of my personal work and that no other
than the indicated aids have been used for its completion. Furthermore I assure
that all quotations and statements that have been inferred literally or in a
general manner from published or unpublished writings are marked as such.
Beyond this I assure that the work has not been used, neither completely nor in
parts, to pass any previous examination.

\vspace{20mm}

\noindent \ThePlace{}, \TheDate{}

\begin{flushright}
  $\overline{~~~~~~~~~~~~~\raisebox{0mm}[5mm]{\TheAuthor{}}~~~~~~~~~~~~~}$
\end{flushright}

>>>>

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                               Acknowledgments                               %
%=============================================================================%

\chapter{Acknowledgments}

This work has been supported by the European Union's Horizon 2020 project
\emph{Coordinated Research in Earth Systems and Climate: Experiments,
  kNowledge, Dissemination and Outreach (CRESCENDO)} under grant agreement
number 641816, the European Union's Horizon 2020 project \emph{Climate-Carbon
  Interactions in the Coming Century (4C)} under grant agreement number 821003,
the European Research Council Synergy Grant \emph{Understanding and Modelling
  the Earth System with Machine Learning (USMILE)} under grant agreement number
855187 and the \emph{Advanced Earth System Model Evaluation for CMIP
  (EVal4CMIP)} project funded by the Helmholtz Society. I acknowledge the
\ac{WCRP}'s \ac{WGCM}, which promoted \acs{CMIP}5 and \acs{CMIP}6. I thank the
climate modeling groups for producing and making available their model output,
the \ac{ESGF} for archiving the data and providing access and the multiple
funding agencies who support \ac{CMIP} and the \ac{ESGF}. The computational
resources of the Deutsches Klima Rechenzentrum (DKRZ, Hamburg, Germany) where
the \ac{ESMValTool} is fully integrated into the \ac{ESGF} infrastructure are
kindly acknowledged.

Personal acknowledgments...
